---
title: "Homework 2"
author: Yiyao LI
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1

Read the Mr. Trashwheel dataset.

```{r, warning=FALSE}
trashwheel_df = 
  read_xlsx(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "Mr. Trash Wheel",
    range = cell_cols("A:N")) %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

Read precipitation data for 2017 and 2018.

```{r}
precip_2018 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
  ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2018) %>%
  relocate(year)

precip_2017 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2017 Precipitation",
    skip = 1
  ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2017) %>%
  relocate(year)
```

Now combine annual precipitation.

```{r}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df = 
  bind_rows(precip_2018, precip_2017)

left_join(precip_df, month_df, by = "month")
```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trash wheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, including some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include month precipitation data.



## Problem 2

Read and clean the NYC subway dataset.

```{r}
nycsubway_df = 
  read_xlsx("./datanyc/NYC_Transit_Subway_Entrance_And_Exit_Data.xlsx") %>%
  janitor::clean_names()
```

```{r}
subway_df = 
  select(nycsubway_df, line:entry, vending, ada) %>%
  mutate(entry = recode(entry, "YES" = "TRUE", "NO" = "FALSE"), entry = as.logical(entry),route8 = as.character(route8), route9 = as.character(route9), route10 = as.character(route10),route11 = as.character(route11))
```

the dataset contains `r names(subway_df)`.

the data cleaning steps contain janitor, select and recode.

the dimension of the resulting dataset is `r nrow(subway_df)`*`r ncol(subway_df)`. 


How many distinct stations are there? 
```{r}
nrow(distinct(subway_df,station_name, line, .keep_all = T))
```

How many stations are ADA compliant?
```{r}
nrow(filter(distinct(subway_df,station_name, line, .keep_all = T), ada == TRUE))
```

What proportion of station entrances / exits without vending allow entrance?
```{r}
nrow(filter(subway_df, entry == FALSE))/nrow(subway_df)
```

Reformat data

```{r}
tidy_subway = pivot_longer(subway_df, route1:route11,
                           names_to = "route_name",
                           values_to = "route_number")
A_train = 
  filter(tidy_subway, route_number =="A") %>%
  distinct(station_name, line)

A_train = 
  filter(tidy_subway, route_number =="A") %>%
  filter(ada == "TRUE") %>%
  distinct(station_name, line)
```

## Problem 3

First step

```{r}
pols_month_df = 
  read.csv("./fivethirtyeight_datasets/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon, c("year", "month", "day")) %>% 
  mutate(president = case_when(prez_gop == "1" ~ "gop", prez_dem == "1" ~"dem")) %>% 
  mutate(month = as.double(month) , month= month.abb[month]) %>% 
  select(-day,-prez_dem,-prez_gop)
```

Second step

```{r}
snp_df = 
  read.csv("./fivethirtyeight_datasets/snp.csv") %>%
  janitor::clean_names() %>% 
  separate(date, c("month", "day", "year")) %>% 
  mutate(month = as.double(month) , month= month.abb[month]) %>%
  relocate(year,month) %>% 
  select(-day)
```

Third step

```{r}
unemployment_df = 
  read.csv("./fivethirtyeight_datasets/unemployment.csv") %>% 
  pivot_longer (Jan:Dec, names_to = "month", values_to = "unemployment") %>% 
  janitor::clean_names() %>% 
  mutate(year=as.character(year))
```

```{r}
pols_snp_unemployment= 
  left_join(pols_month_df,snp_df, by= c("year" = "year","month" = "month")) %>% 
  left_join(unemployment_df, by= c("year" = "year","month" = "month"))
```

